\documentclass[letterpaper,12pt]{article}

\usepackage{geometry, pslatex, fancyhdr, graphicx}
\usepackage{amsmath,amsthm,amssymb,scrextend}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage[makeroom]{cancel}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{mdframed}

\geometry{ margin = 1.0in }

%%% TODO modify these variables as per your homework %%%
\def\myname{Harshit Jain}
\def\myuserid{hmj5262}
%%%%

\pagestyle{fancy}
\lhead{{\bf CMPSC 464 Spring 2024}}
\chead{{\bf Extra Credit Assignment}}
\rhead{{\bf \today}}
\let\newproof\proof
\renewenvironment{proof}{\begin{addmargin}[1em]{0em}\begin{newproof}}{\end{newproof}\end{addmargin}\qed}

\newcounter{problemid}
\stepcounter{problemid}
\def\newproblem{\clearpage\newpage{\bf Problem~\arabic{problemid}\stepcounter{problemid}}\hfill\par}

\setlength\parindent{0em} 
\setlength\parskip{8pt}
\setlength{\fboxsep}{6pt}
 

\begin{document}

\framebox[\textwidth]{
	\parbox{0.96\textwidth}{
		\parbox{0.12\textwidth}{\bf Name:}\parbox{0.6\textwidth}{\myname}\\
		\parbox{0.12\textwidth}{\bf User ID:}\parbox{0.6\textwidth}{\myuserid}
	}
}

%% your solutions %%%


% PROBLEM 1
\newproblem

\begin{algorithm}
    \caption{Checking for Infinite Strings in a DFA}
    \begin{algorithmic}
    \STATE \textbf{Function} \textsc{HasInfiniteStrings}(DFA):
        \STATE visited\_states $\gets$ empty set
        \STATE stack $\gets$ [initial\_state]
    
        \WHILE{stack \textbf{is not} empty}
            \STATE current\_state $\gets$ pop(stack)
            \STATE \textbf{Add} current\_state \textbf{to} visited\_states
    
            \FOR{each transition (current\_state, input\_symbol, next\_state) \textbf{in} DFA}
                \IF{next\_state \textbf{is in} visited\_states}
                    \RETURN \textbf{true}
                \ELSIF{next\_state \textbf{is not in} stack}
                    \STATE push next\_state \textbf{to} stack
                \ENDIF
            \ENDFOR
        \ENDWHILE
    
        \RETURN \textbf{false}
    \end{algorithmic}
\end{algorithm}

The function definition for \textsc{HasInfiniteStrings} takes a DFA as input and returns a boolean value indicating whether the DFA accepts an infinite number of strings.
We initialize an empty set named "visited\_states" to keep track of visited states during the DFS traversal.
We also initialize a stack named "stack" with the initial state of the DFA.
The algorithm enters a while loop that continues until the stack is empty. In each iteration of the loop, it pops the top state (current\_state) from the stack.
The current state is then added to the visited\_states set to mark it as visited.
For each transition originating from the current state in the DFA, the algorithm checks whether the next state (next\_state) has already been visited. If it has been visited, that means there is a loop, and the function returns \underline{True} indicating that the DFA accepts an infinite number of strings. If the next state has not been visited, it is pushed onto the stack for further exploration.
If the while loop completes without finding any loop, the function returns \underline{false}, indicating that the DFA accepts a finite number of strings.

\end{document}